---
title: "HPLC SeaBass Submission"
author: "Sebastian Di Geronimo"
date: '2023-03-13'
format: html
---
# 1.0 Information
## 1.1 Links: SeaBass Instructions for Submission of HPLC
Instructions: https://seabass.gsfc.nasa.gov/wiki/Data_Submission
Check before submit: https://seabass.gsfc.nasa.gov/wiki/FCHECK#Download%20Source%20Code

## 1.2 Header
https://seabass.gsfc.nasa.gov/wiki/metadataheaders

### 1.2.1 Required Fields:
investigators
affiliations
contact
experiment
cruise
station
data_file_name
documetns
data_type
calibration_files
start_date
end_date
start_time
end_time
north_latitude
south_latitude
east_longitude
west_longitude
water_depth
missing
delimiter
fields
units

### 1.2.2 Conditionally Required Headers


### 1.2.3 HPLC Specific
https://seabass.gsfc.nasa.gov/wiki/data_submission_special_requirements#Pigments,%20HPLC
/HPLC_lab (e.g., NASA_GSFC)
/HPLC_lab_technician (e.g., Crystal_Thomas)

### 1.2.4 Ex Header
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ start ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/begin_header	
/identifier_product_doi=10.5067/SeaBASS/EXPORTS/DATA001	
/received=20190807	
/affiliations=Bowdoin_College	
/investigators=Collin_Roesler	
/contact=croesler@bowdoin.edu	sdrapeau@bowdoin.edu
/experiment=EXPORTS	
/cruise=EXPORTSNP	
/data_file_name=EXPORTS_EXPORTSNP_HPLC-inline_survey_20180814.csv	
/original_file_name=EXPORTS_Roesler_08-06_report.xlsx	
/data_type=pigment	
/start_date=20180811	
/end_date=20180908	
/start_time=19:19:00[GMT]	
/end_time=19:44:00[GMT]	
/water_depth=na	
/measurement_depth=5.5	
/west_longitude=-145.960[DEG]	
/east_longitude=-131.543[DEG]	
/north_latitude=50.643[DEG]	
/south_latitude=49.4265[DEG]	
/documents=EXPORTS_Method_Inline_HPLC_Roesler_SeaBASS.docx	EXPORTS_2018_DATA_MASTER.xlsx
/calibration_files=EXPORTS_Method_Inline_HPLC_Roesler_SeaBASS.docx	
/data_status=final	
/missing=-9999	
/below_detection_limit=-8888	
/HPLC_lab=NASA_GSFC	
/HPLC_lab_technician=Crystal_Thomas	
/delimiter=comma	
/fields=hplc_gsfc_id	sample
/units=none	none
/end_header	

~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ end ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

## 1.3 Data fields
1. Individual and summed pigments (based on SeaBASS fields/SeaHARRE reports)

2. Separate columns for any size-fractionated measurements (e.g., Tot_Chl_a_20umprefilt goes in a separate column from Tot_Chl_a)

3. Include "hplc_gsfc_id" as a column if your data were analyzed at the NASA GSFC lab to preserve the lab's sample ID

Report any replicate filters separately
Use the /below_detection_limit to mask any relevant values. It is distinct from the /missing value (e.g., /missing=-9999 vs. /below_detection_limit=-8888)
SeaBASS files may only contain data from a single cruise or deployment. If necessary, sort and separate your data by cruise then create multiple SeaBASS files
An online tool (HPLC2sb) is optionally available to help convert NASA GSFC HPLC spreadsheets into SeaBASS file format. However, please read its limitations and caveats carefully. https://seabass.gsfc.nasa.gov/hplc2sb/

# 2.0 Start Code:
## Load Libraries
```{r libraries}
librarian::shelf(
  librarian, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
  forcats, lubridate, glue, fs, magrittr, here,
  # broom # optional
  
  # additional
  hms, janitor
)

library("conflicted")

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")

# load consolidated HPLC by cruise
(dir_ls(here("data", "processed", "hplc"),
       regexp = "RData") %>%
  sort(decreasing = TRUE))[1] |>
  load() 
```

## Header Information
```{r defaults}
investor  <- "Frank_Muller-Karger,Enrique_Montes"
affil     <- "University_of_South_Florida,USA"
expri     <- "MBON"
hplc_lab  <- "NASA_GSFC"	
lab_tech  <- "Crystal_Thomas"	
type      <- "pigment"
status    <- "final" # final - not intending to revisit
wtr_dpth  <- stn <- "na"

# add more when needed
fields_extr <- c("hplc_gsfc_id", "sample", "date","station","lon","lat","depth",
                 "water_depth","volfilt")
units_extr <- c("none","none","yyyymmdd","none","degrees","degrees","m","m","L")

# TODO:
contact  <- "__TODO__"
docs     <- "__TODO__"
cal_file <- "__TODO__"
comments <- 
  glue(.sep = "\n",
       "!",
       "! test",
       "!")
```

```{r headr-skeleton}
header_skeleton <-
  glue::glue(
    "/begin_header",
    # required  
    "/investigators={investor}",
    "/affiliations={affil}",
    "/contact={contact}",
    "/experiment={expri}",
    "/cruise={cruise_id}",
    "/station={stn}",
    "/data_file_name={file_name}", # current name of data file, self-reference
    "/documents={docs}",           # additional info about experiment and cruise
    "/data_type={type}",
    "/calibration_files={cal_file}",
    "/start_date={start_date}",
    "/end_date={end_date}",
    "/start_time={start_time}[GMT]",
    "/end_time={end_time}[GMT]",
    "/north_latitude={north_latitude}[DEG]",
    "/south_latitude={south_latitude}[DEG]",
    "/east_longitude={east_longitude}[DEG]",
    "/west_longitude={west_longitude}[DEG]",
    "/water_depth={wtr_dpth}",
    "/missing=-9999",
    "/delimiter=comma",
    "/fields={fields}", # fields of data 
    "/units={pun}",  # units for each column 
    
    # HPLC required
    "/HPLC_lab={hplc_lab}",
    "/HPLC_lab_technician={lab_tech}",
    "/below_detection_limit=-8888",
    
    # conditional 
    # "/measurement_depth=___", # only need if don't supply depth field
    
    # optional recommened
    "/data_status={status}",
    
    # optional
    # "/identifier_product_doi=", # entered by SeaBass Staff
    "/original_file_name={fbase}", 
    
    "{comments}",
    
    #"/received=",
    
    "/end_header", 
    # "{dat}",
    .sep = "\n") |>
  
  expression()


```


## Extract Header Information
```{r header-info}
# TODO: delete ----
# pigment_data$data[[1]] %>%
#         mutate(datetime = date + time) %>%
#       
#         summarise(across(.col = c(datetime, lon, lat), 
#                          .fn = list(min = min, 
#                                     max = max),
#                          na.rm = TRUE,
#                          .names = "{.col}_{.fn}")) %>%
#         rename(
#           "north_latitude" = lat_max,
#           "south_latitude" = lat_min,
#           "east_longitude" = lon_max,
#           "west_longitude" = lon_min,
#         ) %>%
#         mutate(
#           start_date = ymd(as_date(datetime_min)),
#           end_date   = ymd(as_date(datetime_max)),
#           start_time = as_hms(datetime_min),
#           end_time   = as_hms(datetime_max),
#           .keep      = "unused"
#         ) 

# ----
test <-
  pigment_data %>%
  mutate(
    header_info = map(
      .x = data,
      function(x) {
        x %>%
        mutate(datetime = date + time) %>%
        # get min and max of head per cruise header info
        summarise(across(.cols = c(datetime, lon, lat), 
                         # new way using lambda notation
                         .fns = list(min = (\(x) min(x, na.rm = TRUE)),
                                     max = (\(x) max(x, na.rm = TRUE))),
                         # alternate using ~
                         # .fns = list(min = ~ min(.x, na.rm = TRUE), 
                         #             max = ~ max(.x, na.rm = TRUE)),
                         
                         # creates warning because `...` in across is deprecated?
                         # .fns = list(min = min, 
                         #             max = max),
                         # na.rm = TRUE,
                         .names = "{.col}_{.fn}")) %>%

        # split datetime to date and time separately
        mutate(
          start_date = format(ymd(as_date(datetime_min)),"%Y%m%d"),
          end_date   = format(ymd(as_date(datetime_max)),"%Y%m%d"),
          start_time = as_hms(datetime_min),
          end_time   = as_hms(datetime_max),
          
          north_latitude = round(lat_max, 3),
          south_latitude = round(lat_min, 3),
          east_longitude = round(lon_max, 3),
          west_longitude = round(lon_min, 3),
          .keep      = "unused"
        )
      }
    )
  ) 
```
Warning: There was 1 warning in `mutate()`.
ℹ In argument: `across(...)`.
Caused by warning:
! There was 1 warning in `summarise()`.
ℹ In argument: `across(...)`.
Caused by warning:
! The `...` argument of `across()` is deprecated as of dplyr 1.1.0.
Supply arguments directly to `.fns` through an anonymous function instead.

  # Previously
  across(a:b, mean, na.rm = TRUE)

  # Now
  across(a:b, \`(x) mean(x, na.rm = TRUE))
This warning is displayed once every 8 hours.
Call `lifecycle::last_lifecycle_warnings()` to see where this warning was generated.

```{r header-info}
names(test)
# "file"        "base"        "cruise"      "param_name"  "units"       "data"        "header_info"
  glue("here", paste(test$param_name[[1]], collapse  = ","), .sep = ",")
test$data[[1]] %>% janitor::remove_empty(which = "cols") %>%
  as.matrix() 
 

# cruise data header_info cruise_id 
test %>%
  slice(1:2) %>%
  mutate(
    cruise_id = case_when(
      str_detect(cruise, "(?i)walton") ~ str_c("WS", str_extract(cruise, "[0-9]+")),
      str_detect(cruise, "(?i)sav") ~ str_c("SV", str_extract(cruise, "[0-9]+")),
      str_detect(cruise, "(?i)weath") ~ str_c("WB", str_extract(cruise, "[0-9]+")),
      TRUE ~ "no"
    ),
     heads = pmap(
      .progress = TRUE,
      .l = list(data, header_info, cruise_id),
      function(y, x, cruise_id){
        cat("\n---\nNEXT\n", first(y$base), "\n---\n")
        
        fbase <- str_replace_all(first(y$base), " ", "_")
        
        file_name <-
          # glue("{expri}", "{cruise_id}", "{type}", "{st_dt}.sb", .sep = "_")
          glue("{expri}", "{cruise_id}", "{type}", "{x$start_date}.sb", .sep = "_")
        
        # fields and units standards: https://seabass.gsfc.nasa.gov/wiki/stdfields
        fields <- glue(!!!fields_extr,
                       paste(first(y$param_name), collapse = ","), .sep = ",")
        
        pun <- glue(!!!units_extr,
                    paste(first(y$units), collapse = ","), .sep = ",")
        
        # cat(file_name, fields, pun, sep = "\n\n")
        cat(x$start_date, "\n", x$end_date, "\n\n\n")
        with(x, eval(header_skeleton)) %T>% print()
        readline("Ready?")
      }
     )
  )
    
  #   (
  #   heads = pmap(
  #     # .progress = TRUE,
  #     .l = list(base, header_info, cruise_id, data, param_name, units, base),
  #     function(files, x, y, dat, names, p_unit, fbase) {
  #       
  #       cat("\n---\nNEXT\n", files, "\n---\n")
  # 
  #       temp   <- x
  #       no_lat <- round(x$north_latitude, 3)
  #       so_lat <- round(x$south_latitude, 3)
  #       es_lon <- round(x$east_longitude, 3)
  #       ws_lon <- round(x$west_longitude, 3)
  #       st_t   <- x$start_time
  #       end_t  <- x$end_time
  #       st_dt  <- format(x$start_date, "%Y%m%d")
  #       end_dt <- format(x$end_date, "%Y%m%d")
  #       cruise_id <- y
  #       fbase <- str_replace_all(fbase, " ", "_")
  # 
  #       file_name <-
  #       glue("{expri}", "{cruise_id}", "{type}", "{st_dt}.sb", .sep = "_")
  # 
  #       fields and units standards: https://seabass.gsfc.nasa.gov/wiki/stdfields
  #       fields <- glue(!!!fields_extr,
  #                      paste(names, collapse = ","), .sep = ",")
  # 
  #       pun <- glue(!!!units_extr,
  #                       paste(p_unit, collapse = ","), .sep = ",")
  # 
  # 
  #      
  #       
  #       
  #       eval(header_skeleton) %T>% print()
  #       readline("Ready?")
  #     }
  #   )
  # ) 


 # dat <- 
 #          dat %>%
 #          remove_empty(which = "cols") %>%
 #          mutate(
 #          volfilt = volfilt/1000 # convert mL to L
 #        ) %>%
 #          as.matrix()

```
# 3.0 Tests
## List Version Test
```{r test-default}
hd_def <- list()
hd_def$invest  <- "Frank_Muller-Karger,Enrique_Montes"

hd_def$affil     <- "University_of_South_Florida,USA"
hd_def$expri     <- "MBON"
hd_def$hplc_lab  <- "NASA_GSFC"	
hd_def$lab_tech  <- "Crystal_Thomas"	
hd_def$type      <- "pigment"
hd_def$status    <- "final" # final - not intending to revisit
hd_def$wtr_dpth  <- hd_def$stn <- "na"
hd_def$comments <- 
  glue(.sep = "\n",
       "!",
       "! test",
       "!")


hd_def$fields_extr <- c("hplc_gsfc_id", "sample", "date","station","lon","lat","depth",
                 "water_depth","volfilt")
hd_def$units_extr <- c("none","none","yyyymmdd","none","degrees","degrees","m","m","L")

# TODO:
hd_def$contact  <- "__TODO__"
hd_def$docs     <- "__TODO__"
hd_def$cal_file <- "__TODO__"


```

```{r test-headr-skeleton}
# "invest"      "affil"       "expri"       "hplc_lab"    "lab_tech"    "type"        
# "status"      "wtr_dpth"    "comments"    "fields_extr"
# "units_extr"  "contact"     "docs"        "cal_file"   

header_skeleton <-
  glue::glue(
    "/begin_header",
    # required  
    "/investigators={hd_def$invest}",
    "/affiliations={hd_def$affil}",
    "/contact={hd_def$contact}",
    "/experiment={hd_def$expri}",
    "/cruise={cruise_id}",
    "/station={hd_def$stn}",
    "/data_file_name={file_name}", # current name of data file, self-reference
    "/documents={hd_def$docs}",           # additional info about experiment and cruise
    "/data_type={hd_def$type}",
    "/calibration_files={hd_def$cal_file}",
    "/start_date={st_dt}",
    "/end_date={end_dt}",
    "/start_time={st_t}[GMT]",
    "/end_time={end_t}[GMT]",
    "/north_latitude={no_lat}[DEG]",
    "/south_latitude={so_lat}[DEG]",
    "/east_longitude={es_lon}[DEG]",
    "/west_longitude={ws_lon}[DEG]",
    "/water_depth={hd_def$wtr_dpth}",
    "/missing=-9999",
    "/delimiter=comma",
    "/fields={fields}", # fields of data 
    "/units={pun}",  # units for each column 
    
    # HPLC required
    "/HPLC_lab={hd_def$hplc_lab}",
    "/HPLC_lab_technician={hd_def$lab_tech}",
    "/below_detection_limit=-8888",
    
    # conditional 
    # "/measurement_depth=___", # only need if don't supply depth field
    
    # optional recommened
    "/data_status={hd_def$status}",
    
    # optional
    "/original_file_name={fbase}", 
    
    "{comments}",
    
    #"/received=",
    
    "/end_header", 
    # "{dat}",
    .sep = "\n") |>
  
  expression()



# header_skeleton
# eval(header_skeleton)
```

```{r}
# for (i in 1:2) {
#   cat("\n---\nNEXT\n", test$file[[1]], "\n---\n")
#   temp   <- test$header_info[[i]]
#   no_lat <- round(temp$north_latitude, 3)
#   so_lat <- round(temp$south_latitude, 3)
#   es_lon <- round(temp$east_longitude, 3)
#   ws_lon <- round(temp$west_longitude, 3)
#   st_t   <- temp$start_time
#   end_t  <- temp$end_time
#   st_dt  <- format(temp$start_date, "%Y%m%d")
#   end_dt <- format(temp$end_date, "%Y%m%d")
#   eval(header_skeleton) %T>% print()
#   
# }
```
x  hdr = {'/begin_header';
  strcat('/received=',date_received);
  '/investigators=Frank_Muller-Karger,Enrique_Montes';
  '/affiliations=University_of_South_Florida,USA';
'/contact=emontesh@usf.edu';
'/experiment=MBON';
strcat('/cruise=',cruise);
strcat('/data_file_name=',cruise,'_',data_type);
strcat('/documents=',cruise_report);
'/calibration_files=HPLC_method_summary';
'/delimiter=space';
'/data_type=pigment';
'/data_status=final';
strcat('/start_date=',start_date);
strcat('/end_date=',end_date);
strcat('/start_time=',start_time,'[GMT]');
strcat('/end_time=',end_time,'[GMT]');
strcat('/north_latitude=',lat_n,'[DEG]');
strcat('/south_latitude=',lat_s,'[DEG]');
strcat('/west_longitude=',lon_w,'[DEG]');
strcat('/east_longitude=',lon_e,'[DEG]');
'/below_detection_limit=-8888';
'/missing=-8888';
'/water_depth=-999';
'/cloud_percent=-999';
'/wave_height=-999';
'/wind_speed=-999';
'/secchi_depth=-999';
'!';
'!RUN BY: Enrique Montes';
'!Samples were processed at NASA GSFC by Crystal Thomas';
'!Measurements below detection limit are assigned the value -9999; this represents pigments not detected';
'!as well as pigments present but with concentrations less than 0.0005?g/L.';
'!Coefficient of variation (replicate filter precision) for TChl a =	1.64%; Ppig =	3.74%';
'!!Sky: variable conditions';
'!Water: variable conditions.';
'/fields=date,station,lon,lat,depth,Tot_Chl_a,Tot_Chl_b,Tot_Chl_c,alpha-beta-Car,But-fuco,Hex-fuco,Allo,Diadino,Diato,Fuco,Perid,Zea,MV_Chl_a,DV_Chl_a,Chlide_a,MV_Chl_b,DV_Chl_b,Chl_c1c2,Chl_c3,Lut,Neo,Viola,Phytin_a,Phide_a,Pras,Gyro,TChl,PPC,PSC,PSP,Tcar,Tacc,Tpg,DP,Tacc_TChla,PSC_Tcar,PPC_Tcar,TChl_Tcar,PPC_Tpg,PSP_Tpg,TChla_Tpg';
'/units=yyyymmdd,none,degrees,degrees,m,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,mg/m^3,none,none,none,none,none,none,none';
'/end_header'};