---
title: "Add Depth Level"
author: "Sebastian Di Geronimo"
format: html
editor: source
---

WIP: modify to be more generalized
```{r setup}
librarian::shelf(
  librarian, 
  tidyverse, here, readxl, fs, magrittr, janitor, glue
)
```

```{r}
# rstudioapi::selectFile()

dat <- 
  here("C:/Users", "imars_cruise", "Documents",
       "s_fl_cruise_data", 
     "old_cruises_for_reference",
     "imars_inventory_chl_hplc_cdom_update_feb_06_2023.xlsx") %>% 
  read_xlsx(sheet = 6) %>%
  arrange(date_mm_dd_yy) 

dat %>% 
  count(
    year, cruise_id, station
  ) %>% 
  filter(cruise_id != "WS18285") %>% 
  arrange(desc(n))
```
```{r}
depths <- c("Surface", "Mid", "Bottom")

dat_rank <- dat %>% 
  distinct(cruise_id, year, date_mm_dd_yy, station, depth_m) %>% 
  # filter(is.na(depth_m))
  mutate(
    .after = depth_m,
    .by = c(cruise_id, station),
    depth_m = if_else(is.na(depth_m), 1, depth_m),
    depth_rank = dense_rank(depth_m),
    depth_lvl = case_when(
      n() == 1 & depth_m < 6 ~ depths[1], # surface
      n() == 1 & depth_m >= 6 ~ depths[3], # bottom
      n() == 2 & depth_rank == 1 & depth_m < 6 ~ depths[1], # surface
      n() == 2 & depth_rank == 2 & depth_m >= 5 ~ depths[3], # bottom
      n() == 2 & depth_rank == 2 & depth_m >= 4.3 & str_detect(station, "MR") ~ depths[3], # bottom
      n() == 3 & depth_rank == 2 ~ depths[2], # "Mid",
      n() == 3 & depth_rank == 1 ~ depths[1], # "surf",
      n() == 3 & depth_rank == 3 ~ depths[3], #"bott",
      .default = "idk"
    )
  ) %T>%
  print()
```


```{r}
# check handwritten logs or NOAA log sheet for these cruises and stations
dat_rank %>% 
  filter(.by = c(cruise_id, station),
         any(str_detect(depth_lvl, "idk"))
         ) %>%
  View()
```

Notes:
SV18067 
  - check WS, depths are 2 and 4.8 (probably 48?)
SV18173 
  - check MR, depths are 1.6 and 4.4 (probably 44?)
WS18285 
  - check TB6 depth, one is 2 other is 4, checks out from NOAA log
WS18351
  - check 7, depths are 1.9 and 1.5
WS19210 
  - check 68, depth are 2 and 2.4,
  - check dates, they are different too, Jul 30 and Aug 2
  - check GP1, depths are 1.84 and 3.32 - checks out from NOAA log
WS21277
  - check MR, depths are 1 and 4.3, probably bottom was NA and marked as 1

```{r}
dat %>% 
  # two samples issues with duplication
  filter(!str_detect(identifier, "WS23011-180|SV18067-CK")) %>% 
  # one cruise issues with identifier duplication
  filter(!str_detect(base, "(?i)aug2015")) %>% 
  left_join(dat_rank) %>% 
  janitor::get_dupes(identifier)
```


# Create a Lists of Samples Collected to Send to NOAA
They have requested we create a spreadsheet with sample station and qualitative
depth (surface, mid, bottom) for and combine all sample types collected with an
"x". 

We collect chl-a spectra, HPLC, CDOM and eDNA on every cruise so at least these
many columns are needed. BB3 and Rrs may be included if taken on this cruise.

## Cruise File Path
```{r}
cruise_path <- 
  here("C:/Users", "imars_cruise", "Documents", "s_fl_cruise_data") %>% 
  dir_ls(regexp  = "fknms",
         recurse = 2) %>% 
  str_subset("~", negate = TRUE) %T>% 
  print()
cruise_path <- 
  here("C:/Users", "imars_cruise", "Documents", "s_fl_cruise_data") %>% 
  rstudioapi::selectFile(path = .) %T>% 
  print()

if (length(cruise_path) > 1) {
  cruise_path <- cruise_path[menu(basename(cruise_path))]
}
```


## Chla, HPLC, CDOM samples
```{r}
sample_collect <-
  (readxl::excel_sheets(cruise_path)) %>% 
  str_subset("field") %>% 
  read_excel(
    path  = cruise_path,
    sheet = .,
    .name_repair = janitor::make_clean_names
  ) %>% 
  # filter(!is.na(station)) %>%
  mutate(temp_depth = if_else(is.na(depth_m), 0, depth_m)) %>%
  filter(!is.na(vol_ml)) %>% 
# --- probably not needed
  # filter(
    # .by = c(station, temp_depth),
    # !any(is.na(date_mm_dd_yy))
    # & !any(is.na(sample_collection_time_gmt))
  # )
# ---

  mutate(
    .before   = 1,
    collected = str_extract(identifier, "\\w{1,2}\\d{4,5}"),
    collected = glue::glue("{collected}-{station}-{depth_level}"),
    collected = str_replace(collected, "LK", "LK/21"),
    x         = "x"
  ) %>% 
  # --- may remove if don't want number of samples taken per type
  # i.e. chl-a "x (2)" vs "x"
  # mutate(
  #   .by = c(station, depth_m, sample_type),
  #   x   = glue::glue("{x} ({n()})"),
  #   x   = str_remove(x, " \\(1\\)")
  # ) %>%
  # ---
  distinct(collected, x, station, depth_m, sample_type) %>% 
  pivot_wider(
    names_from  = sample_type,
    values_from = x
  ) %>% 
  select(collected, last_col(2):last_col()) %T>%
  print()
```


## eDNA
```{r}
edna_collect <-
  (readxl::excel_sheets(cruise_path)) %>% 
  str_subset("(?i)edna") %>% 
  read_excel(
    path  = cruise_path,
    sheet = .,
    .name_repair = janitor::make_clean_names
  ) %>% 
  select(1:contains("note")) %>% 
  filter(!is.na(vol_ml)) %>% 
  mutate(sample_type = "eDNA") %>%
  # filter(!is.na(station) & !str_detect(station, "0")) %>%
  # mutate(
  #   temp_depth  = if_else(is.na(depth_m), 0, depth_m),
  #   sample_type = "eDNA"
  # ) %>%
  # filter(
  #   .by = c(station, temp_depth),
  #   !any(is.na(mm_dd_yy)) 
  #   & !any(is.na(time_gmt))
  # ) %>% 
  mutate(
    .before     = 1,
    # depth_level = if_else(
    #   str_detect(depth_level, "0"),
    #   NA,
    #   depth_level),
    collected   = str_extract(sample, "\\w{1,2}\\d{4,5}"),
    collected   = glue::glue("{collected}-{station}-{depth_level}", .na = ""),
    collected   = str_replace(collected, "LK", "LK/21"),
    x = "x"
  ) %>% 
  # --- may remove if don't want number of samples taken per type
  # mutate(
  #   .by = c(station, depth_m),
  #   x   = glue::glue("{x} ({n()})", .na = ""),
  #   x   = str_remove(x, " \\(1\\)")
  # ) %>% 
  # ---
  distinct(collected, x, station, depth_m, sample_type) %>% 
  pivot_wider(
    names_from  = sample_type,
    values_from = x
  ) %>% 
  select(collected, last_col()) %T>%
  print()
```

## Join Both Sheets
```{r}
both_sheets <- 
  full_join(
    sample_collect,
    edna_collect
    ) %T>%
  print()
```


# Save collected samples
```{r save-collected}
overwrite <- 0
cruise_id <-
  cruise_path %>% 
  str_extract("\\w{1,2}\\d{4,5}")

file_name <-
  here("C:/Users", "imars_cruise", "Documents",
       "s_fl_cruise_data", 
        cruise_id, "metadata", 
        glue("imars_samples_collected_{cruise_id}.xlsx") )

if (!file_exists(file_name)) {
  openxlsx::write.xlsx(
    x         = both_sheets, 
    file      = file_name, 
    overwrite = FALSE
    )

  shell.exec(file_name)
  
} else {
  message("File already exists.")
  cli::cli_alert_warning("Do you want to overwrite?\n")
  overwrite <- 
    menu(c("Yes", "No"), title = "[Select 1 = Yes or 2 = No in console]")
}

if (overwrite == 1) {
  cli::cli_alert_info("Overwriting Previous File!")
  openxlsx::write.xlsx(
    x = both_sheets, 
    file = file_name, 
    overwrite = TRUE)
  
  shell.exec(file_name)
}

```

