---
title: "Min and Max Date per Cruise"
author: "Sebastian DiGeronimo"
date: "2023-03-20"
output: html_document
---
# Info

This script will create a fill with cruise ID, min date and max date. This data
used here is from NOAA AOML. It is summarized per cruise for the min data and
max date. This is used to verify the date written on zooplankton cod ends with a 
cruise ID when not written. This info will be used to print labels to be placed
on vials when transfering zooplankton in formalin to zooplankton in Ethanol.

Errors found:
- WS18120 has two different months
  - WS18120 is before Aug 2018 and WS18218 is after Aug 2018
- WS17170 
  - had multiple names:
     WS17171, WS17172, WS17173
- WS17030
  - had half of cruise with the year `2016`

Modified info to match IMaRS:
- SAV (NOAA) vs SV (IMaRS) in cruise ID
- SV1803 (NOAA) vs SV18067 (IMaRS) 
- WS20231 (NOAA) vs WS20230 (IMaRS)


# Load Libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
librarian::shelf(
  librarian, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
  forcats, lubridate, glue, fs, magrittr, here,
  # broom # optional
  
  # additional
  readxl, openxlsx, janitor
)

library("conflicted")

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
```

# Create Dates File
```{r}
(here("data", "raw") %>%
dir_ls(regexp = "WSMasterSampleLog",
       recurse = TRUE))[1] %>%
read_xlsx(
  guess_max = 7000,
  na = "Skip too shallow",
  .name_repair = make_clean_names
  ) %>%
select(cruise, date_gmt) %>%
mutate(
  cruise   = str_replace(cruise, "SAV", "SV"),
  cruise   = case_when(str_detect(cruise, "WS18120")
                       & date_gmt > as_date("2018-08-01") ~ "WS18218",
                       str_detect(cruise, "WS20231")      ~ "WS20230",
                       str_detect(cruise, "SV1803")       ~ "SV18067",
                       str_detect(cruise, "WS1717")       ~ "WS17170",                   
                       .default = cruise),
  
  date_gmt = if_else(year(date_gmt) < 2017 
                     & str_detect(cruise, "WS17030"),
                     as_date(str_replace(date_gmt, "2016", " 2017")),
                     date_gmt),
  .keep    = "used"
) %>%
summarise(.by = cruise,
          min_date = min(date_gmt, na.rm = TRUE),
          max_date = max(date_gmt, na.rm = TRUE),
          ) %>%
arrange(min_date) %>%
mutate(
  min_date = format(min_date, "%m/%d/%Y"),
  max_date = format(max_date, "%m/%d/%Y"),
) %>%
write_csv(file = here("data", "metadata","cruise_date.csv"))
```