---
title: "Merge Incubations with Metadata"
author: "Sebastian Di Geronimo"
date: "2023-10-18"
format: html
---

Work in progress

# 1.0 ---- Summary of Document ----

This script is used to merge data and metadata from O2 measurements to estimate
primary production. 

This data comes from NOAA AOML as part of the South Florida Ecosystem 
Restoration project (SFER) in partnership with MBON.


--
NOTE: 
--
This script was made quickly so it may not work across computers

# 2.0 ---- Setup ----

## 2.1 Load Libraries
```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
  install.packages("librarian")

librarian::shelf(
  quiet = TRUE,
  librarian, conflicted, dplyr, stringr,
  forcats, lubridate, fs, magrittr, here,
  
  # additional
  openxlsx, janitor, rstudioapi
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
  )

```

## 3.0 Set Dataset Path

```{r load-data}
# file_incubation <-
#   here("data", "raw") %>%
#   dir_ls(regexp = "Incubations\\.xlsx") %T>% print()
 
# file_metadata <-
#   here("data", "raw") %>%
#   dir_ls(regexp = "SampleLog") %>%
#   str_subset("~", negate = TRUE) %T>% print()

file_incubation <- rstudioapi::selectFile(caption = "Select Incubation File") %T>% print()
file_metadata <-  rstudioapi::selectFile(caption = "Select Metadata File") %T>% print()
```

# 4.0 Load Datasets

```{r load-data}
# ---- metadata
dat_meta <- 
  openxlsx::read.xlsx(
  file_metadata,
  sheet = "All Depths",
  detectDates = TRUE) %>%
  janitor::clean_names(replace = janitor:::mu_to_u) %>%
  select(1, cruise:depth)  %>%
  relocate(station, .after = cruise)  %>%
  
  # fix station and cruise names
  mutate(
    station = str_replace(station, "^LK", "21/LK"),
    station = str_replace(station, "^21$", "21/LK"),
    station = str_replace(station, "^21.5/LK$", "21/LK"),
    station = case_when(
      rank == 94 ~ "54b",
      rank == 91 ~ "57b",
      .default = station
    ),
    cruise = if_else(
      between(date_gmt, as_date("2018-07-01"), as_date("2018-08-30")),
      "WS18218",
      cruise)
  )  %>%
  
  # remove bad stations
  filter(
    rank != 63
    & rank != 914
    & rank != 1165
  ) %T>% print()

# ---- incubation
dat_incubation <- 
  openxlsx::read.xlsx(
  file_incubation,
  sheet = "Final"
  ) %>%
  janitor::clean_names(replace = janitor:::mu_to_u)  %>%
  
  # fix station and cruise names
  mutate(
    station = str_replace(station, "^LK", "21/LK"),
    cruise = str_replace(cruise, "SAV18069", "SAV1803")
    ) %T>% print()
```


# 5.0 Get Distinct Metadata with Incubations as a Reference

```{r}
# ---- filter incubation to distinct cruise and station
filt_incubation <- 
  dat_incubation %>%
  distinct(cruise, station) %T>% print()


# ---- filter metadata to distinct cruise and station
filt_meta <- 
  dat_meta %>%
  select(-rank) %>%
  arrange(cruise, station, desc(depth)) %>%
  distinct(cruise, station, date_gmt, .keep_all = TRUE) %T>% print()

# ---- merge filtered metadata and filtered incubation data
full_metadata <- 
  inner_join(
    filt_meta,
    filt_incubation, 
    by = c("cruise", "station")
    ) %T>% print()

# ---- show missing incubations
anti_join(
  filt_incubation, 
  dat_meta,
  by = c("cruise", "station")
  )
```

# 6.0 Save Merged Incubations and Metadata 

```{r}
save_loc <-
  rstudioapi::selectDirectory(caption = "Select Save Location") %>%
  here("merged_incubations_metadata.xlsx") %T>% print()


dat_incubation %>%
  mutate(
    .before = 1,
    rank    = fct_inorder(as.character(row_number()))
  ) %>%
  right_join(
    full_metadata,
    .,
    by = c("cruise", "station")
  ) %>%
  arrange(rank) %T>% print() %>%
  openxlsx::write.xlsx(
    file      = save_loc,
    overwrite = FALSE
  )


shell.exec(save_loc)

```



