---
title: "Convert eDNA Metadata to NOAA Format"
author: "Sebastian Di Geronimo"
date: "2024-03-13"
format: html
editor: source
---

# 1.0 ---- Summary of Document ----

This document takes the metadata from the cruise logsheet and converts it into
the NOAA preferred format. This is to reduce additional steps taken by the user
at sea to make a seamless transition.


# 2.0 ---- Load Libraries ----

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
  install.packages("librarian")

librarian::shelf(
  quiet = TRUE,
  librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
  forcats, lubridate, glue, fs, magrittr, here,
  
  # additional
  openxlsx2
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
  )
  
if (!exists("cloud_dir")) source(here(".Rprofile"))
```

# 3.0 ---- eDNA Metadata ----

## 3.1 Find NOAA eDNA Template

The template is used to create the header for the eDNA metadata.

This template is stored in:

<cloud directoryr location on computer>/Box/mbon_imars_cruises/
└── blank_sheets
    └── edna_cruise_template_2024.xlsx <--- *this file*

```{r noaa-edna-template}
noaa_template <-
  here(cloud_dir, "blank_sheets") %>%
  dir_ls(regexp = "edna_.*template") %>%
  str_subset("~", negate = TRUE) %T>% 
  print()
```

## 3.2 Select Metadata

This will have a popup to select the metadata file that contains a sheet called
`eDNA`. This file name prefix should have `fknms_` and should be in its `cruise` 
folder within *Box*

Example Structure:

<cloud directory location on computer>/Box/mbon_imars_cruises/
└── years
    └── 2024
        └── HG24013
            ├── above_water_radiometry
            ├── apad
            ├── bb3
            ├── CDOM
            ├── forms
            └── metadata
                ├── fknms_sample_logsheet_03_2024_HG24062.xlsx  <--- *this file*
                ├── HG24062_edna_sample_locations.jpeg
                ├── HG24062_imars_sample_locations.jpeg
                └── HG24062_imars_sample_locations_collected.jpeg

```{r select-metadata}
# metadata
metadata <-
  rstudioapi::selectFile(path = here(cloud_dir, "years")) %T>%
  print()

# sheet name
sheet_name <- 
  readxl::excel_sheets(metadata) %>%
  str_subset("eDNA$") %T>% 
  {
    glue("Sheet Name: {.}") %>%
    print()
  }
```


## 3.3 Load and Reformat eDNA Metadata

```{r load-format-metadata}
noaa_format <- 
  readxl::read_xlsx(
    metadata, 
    sheet = sheet_name, 
    .name_repair = janitor::make_clean_names
  ) %>%
  janitor::remove_empty(which = "cols") %>%
  filter(!is.na(mm_dd_yy) & !year(mm_dd_yy) < 2000 & !is.na(vol_ml)) %T>% 
  print() %>%
  mutate(
    .after                = time_gmt,
    station               = str_replace(station, "LK", "21/LK"),
    time_gmt              = hms::as_hms(time_gmt),
    date_time             = as_datetime(paste(mm_dd_yy,time_gmt), tz = "UTC"),
    `*collection_date`    = format_ISO8601(date_time, usetz = "Z"),
    date_time_local       = with_tz(date_time, "America/New_York"),
    collection_date_local = format_ISO8601(date_time_local,  usetz = TRUE),
    collection_date_local = str_replace(collection_date_local, "(-\\d{2})00$", "\\1:00")
  )  %T>% 
  print() %>%
  
  select(
  `*sample_name`      = sample_name,
  serial_number       = sample,
  cruise_id,
  line_id,
  line_name,
  station,
  ctd_bottle_no       = niskin_bottle,
  sample_replicate,
  notes_sampling      = notes,
  collection_date_local,
  `*collection_date`,
  `*depth`            = depth_m,
  tot_depth_water_column = max_depth_m,
  depth_category      = depth_level,
  `*geo_loc_name`     = geo_loc_name,
  waterBody           = water_body,
  `*lat_lon`          = lat_lon,
  decimalLatitude     = lat,
  decimalLongitude    = lon,
  samp_vol_we_dna_ext = vol_ml,
  ) %>%
  mutate(
    waterBody = as.character(waterBody),
    across(
      c(where(is.character), waterBody), 
      ~ na_if(.x, "0")
      )
    
  ) %T>% 
  print()
```

# 4.0 ---- Save Metadata ----

## 4.1 Save Locally

Save the created eDNA file locally to look for any errors.

Naming convention: NOAA_eDNA_metadata_<Cruise ID>.xlsx

```{r local-save}
# extract cruise name
cruise_name <-
  str_extract(metadata, "[A-Z]{1,2}\\d{4,5}") %>%
  str_c(collapse = "_")


# create local directory
local_dir <- here("data", "metadata", "edna_meta", "send_noaa")
dir_create(local_dir)

# create file name
file_name <- 
  here(
    local_dir,
    glue("NOAA_eDNA_metadata_{cruise_name}.xlsx")
    )

# load template
wb <- wb_load(noaa_template)
wb_get_sheet_names(wb)

# remove extra sheets
wb <- wb_remove_worksheet(wb, sheet = "Sheet1")
wb <- wb_remove_worksheet(wb, sheet = "Sheet2")

# add data
wb <- wb_add_data(wb, x = noaa_format, sheet = "noaa_format", na.strings = "")
wb_get_sheet_names(wb)

# save workbook and open
wb_save(wb, file_name)
shell.exec(file_name)
```
## 4.2 Copy NOAA eDNA Format to Cloud

If everything looks okay, copy the eDNA metadata to the cruise folder.

```{r copy-cloud}
file_copy(
  file_name,
  dirname(metadata)
)
```

# 5.0 Create Raw eDNA File 

This basically copies the eDNA sheet from the original log sheet to send to 
NOAA for their records

edna_imars_raw_{cruise ID}.xlsx

```{r raw-metadata}
file_name_raw <-
  here(
    dirname(metadata),
    glue(
      "edna_imars_raw_{cruise_name}.xlsx"
    )
  ) %T>% 
  print()

wb_raw <- read_xlsx(metadata, sheet = sheet_name) 

wb_names <- names(wb_raw)[-which(is.na(names(wb_raw)))]

wb_raw <- 
  wb_raw %>%
  janitor::clean_names() %>%
  select(!matches("^na_?")) %>%
  filter(!is.na(vol_ml)) %>%
  mutate(
    time_gmt = hms::as_hms(time_gmt),
    water_body = NA
  ) %>%
  rlang::set_names(wb_names) %T>%
  print()


write_xlsx(wb_raw, file_name_raw, na.strings = "")
```

