---
title: "Sample Process Log Sheet"
author: "Sebastian DiGeronimo"
date: "2023-02-13"
output: html_document
---
# Info:
This script creates process logs sheets within box. These are formatted from the
file imars_inventory_chl_hplc_cdom.xlsx. They are used to fill in
file name and Turner Triology values.





# ---- Load Libraries ----


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!nzchar(system.file(package = "librarian"))) 
  install.packages("librarian")

librarian::shelf(
  quiet = TRUE,
  librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
  forcats, lubridate, glue, fs, magrittr, here,
  
  # additional
  openxlsx, cli
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
  )

source(here("scripts", "process_log_sheet.R"))


sht_nm <- c("metadata", "Chl-a")
```


# ---- Triology Calibration Information ----

This will need to be updated over time
Update: "2022-04-29", 0.17441, -0.501753918982416, 27.32, 200.39
Update: "2023-03-28",
Update: "2024-02-02", 0.17478, -0.27767397, 27.52, 191.97

```{r triology-calibration}
cal <- 
  tribble(
    ~cal_date,    ~no_acid_slop, ~no_acid_y_int,     ~cal_no_acid_low, ~cal_no_acid_high,
    # <date>      <dbl>          <dbl>               <dbl>            <dbl>
    "2022-04-29", 0.17441,      -0.501753918982416, 27.32,            200.39,
    "2023-03-28", 0.177357021947293, 0.134407327675063, 27.52,        192.11,
    "2024-02-02", 0.17478,      -0.27767397,        27.52,            191.97
    # "2024-04-01", 1, 1, 1, 11,

    ) %>%
  mutate(
    cal_date = as_date(cal_date)
  )
cal

# Change to TRUE if need to update calibration for process logs
update_cal <- FALSE 
# update_cal <- TRUE
```



# ---- Read Log and filter Cruise IDs without Process Log File ----



```{r read-log}
# cruise IDs that already have process logs
crs_log_exst <- 
  here(cloud_dir, "process_logs") %>%
  dir_ls() %>%
  tibble(files = .) %>%
  mutate(
    cruise_id = str_extract(files, "[A-Z]{1,3}\\d{4,5}"),
    cruise_id = if_else(is.na(cruise_id), str_extract(files, "[A-Z]{4}_\\d"), cruise_id)
  ) %T>% print()

# cruise IDs without process log
meta_all <-
  here(cloud_dir) %>%
  dir_ls(regexp = "imars_inventory_chl_hplc_cdom") %>%
  str_subset("~", negate = TRUE) %>%
  openxlsx::read.xlsx(sheet = "sample_meta_data") %>%
  mutate(
    date_time     = janitor::convert_to_datetime(date_time, tz = "UTC"),
    date_mm_dd_yy = janitor::convert_to_date(date_mm_dd_yy, tz = "UTC"),
    sample_collection_time_gmt = str_remove(sample_collection_time_gmt, "\\.\\d+"),
    sample_collection_time_gmt = hms::parse_hms(sample_collection_time_gmt),
  ) %>%
  # remove cruises fully completed pre-COVID
  filter(
    year >= 2019 & !str_detect(cruise_id, c("WS19028|WS19119|WS19210"))
    & !(cruise_id %in% crs_log_exst$cruise_id)
    ) %T>% 
  print()

if (update_cal) {
  cli_alert_warning(
    "You've set update_cal = {update_cal}. You will be updating the log!"
  )
} else {
  stopifnot("No files need creating!" = nrow(meta_all) > 1)
}

```

## ---- Create Workbook ----

function: process_log_sheet(.meta, cal, cruise)
Uses logsheet info for data and cruise ID
Uses calibration variable to input into new ones

If a new cruise process log needs to be created, it will be created here.
If the calibration needs to be updated, `update_cal` needs to be set to `TRUE`
and it will update all files that already exists with the new calibration if the
sample has not been run yet.



```{r workbook}
if (!update_cal) {
save_meta <-
  meta_all %>% 
  nest(.by = cruise_id) %>%
  mutate(
    sheets = map2(data, cruise_id,
      ~ tryCatch({
        cli_alert_info(c("-----\n Cruise ID: ", .y))
        process_log_sheet(.meta = .x, cal = cal, cruise = .y)
      }, error = function(e) {
        cli_alert_danger("Cruise has issue unnesting!")
        cli_alert_danger(sprintf("Error in %s: %s",
                                 deparse(e[["call"]]),
                                 e[["message"]])
                              )
        NULL
        })
      )
    )
} else if (update_cal) {
  save_meta <- 
    crs_log_exst %>%
    # if need to udpate a specific cruise
    # filter(str_detect(cruise_id, "<cruise ID>")) %>%
    mutate(
      sheets = map(
        .x = files,
        ~ update_proc_log(.x, cal, type = "cal"),
        .progress = TRUE
      )
    )

} else {
  cli_alert_danger("Stop?")
}
```


## ---- Save New Worksheet or Updated Calibration ----

Prior to running, check arg list to make sure nothing goes wrong

ignore = TRUE   - checks to redo all or ignore some
overwrt = FALSE - skip saving of files if file exist
dir_loc = local - save to a local folder instead of cloud to verify changes

```{r save}
arg <- 
  list(
    ignore  = TRUE, # redo all or ignore some
    # ignore  = FALSE, # redo all or ignore some
    overwrt = FALSE,
    # overwrt = TRUE,
    
    # directory location
    dir_loc = here("data", "process_logs", Sys.Date()) # local
    # dir_loc = here(cloud_dir, "process_logs") # cloud
  )

dir_create(arg$dir_loc)

cli_alert_info("File save location: {.file {dirname(arg$dir_loc)}}")

for (i in seq(nrow(save_meta))) {
  
  if (nrow(meta_all) > 0) {
    arg$file_name <-
      here(
        arg$dir_loc,
        glue(
          "{save_meta$cruise_id[i]}_",
          "{save_meta$data[[i]]$year[[1]]}_",
          "{min(format(save_meta$data[[i]]$date_mm_dd_yy, \"%m\"), na.rm = TRUE)}",
          "_abs.xlsx"
        )
      )
  } else if (update_cal) {
    arg$file_name <-
      here(arg$dir_loc, basename(save_meta$files[[i]]))
  }
  
  # ---- if need to skip a file because currently being used ---- #
  # if (str_detect(arg$file_name, "<cruise ID>")) {
    # cli_alert_warning("Skipping file {.file {basename(arg$file_name)}}")
    # next()
  # }
  
  if (!file_exists(arg$file_name) | !arg$ignore) {
    cli_alert_info("Trying to create file {.file {basename(arg$file_name)}}")
    try(
      saveWorkbook(
        wb          = save_meta$sheets[[i]],
        file        = arg$file_name,
        overwrite   = arg$overwrt,
        returnValue = TRUE
      ),
      silent = FALSE
    )
  } else {
    cli_alert_info("Ignoring file {.file {basename(arg$file_name)}}")
  }
}

```

# Update Metadata

To update a process log file, add the cruise ID to `cruise_to_update` in quotes.
If multiple are needed, separate them with a comma. If need to update all of 
them, add "all" and this will update all files.

```{r}
cruise_to_update <- c("all")

if (any(str_length(cruise_to_update) > 0)) {
  temp <- 
    meta_all %>%
    nest(.by = cruise_id) %>%
    
    { if (!any(str_detect(cruise_to_update, "all"))) {
    filter(., str_detect(cruise_id, str_c(cruise_to_update, collapse = "|"))) 
      } else {.}
    } %>%
    
    left_join(crs_log_exst) %>%
    filter(!is.na(files)) %>%
    mutate(
      sheets = pmap(
        list(data, cruise_id, files),
        function(.x, .y, .z) {
          tryCatch(
            {
              update_proc_log(
                x           = .x, 
                cruise      = .y, 
                prev_log    = .z, 
                type        = "meta",
                .keep_wksht = TRUE
                )
            },
            error = function(e) {
              cli_alert_danger("Cruise has issue unnesting!")
              cli_alert_danger(sprintf(
                "Error in %s: %s",
                deparse(e[["call"]]),
                e[["message"]]
              ))
              NULL
            }
          )
        }
      )
    ) 
  
  if (menu(c("Yes", "No"), title = "Do you want to open before saving?") == 1) {
   map(temp$sheets, \(x) openXL(x))
  }
  
  temp %>%
  with(., (walk2(
    files, sheets,
    (\(.x, .y) {
      cli_alert_info("Updating file {.file {basename(.x)}}")
      shell.exec(dirname(.x))
      try(
        saveWorkbook(.y,
          .x,
          overwrite   = TRUE,
          returnValue = TRUE
        ),
        silent = FALSE
      )
    })
  )
  ))
}
```


```{r}
# cruise_to_update <- c("WS21277")
# meta_all <- 
#   dir_ls(here(cloud_dir), 
#          regexp = "imars_inventory_chl_hplc_cdom") %>%
#   openxlsx::read.xlsx(sheet = "sample_meta_data") %>%
#   # remove cruises fully completed pre-COVID
#   filter(
#     year >= 2019 & !str_detect(cruise_id, c("WS19028|WS19119|WS19210"))
#     # & !(cruise_id %in% crs_log_exst$cruise_id)
#     ) %>%
#     mutate(
#       date_time = janitor::convert_to_datetime(date_time, tz = "UTC"),
#       date_time = as.POSIXct(date_time, tz = "UTC"),
#       date_mm_dd_yy = janitor::convert_to_date(date_mm_dd_yy, tz = "UTC"),
#       
#     ) 
# 
# meta_all %>%
#     nest(.by = cruise_id) %>%
#     filter(str_detect(cruise_id, str_c(cruise_to_update, collapse = "|"))) %>%
#     mutate(files = dir_ls(here(), regexp = "WS212")) %>%
#     
#   # left_join(crs_log_exst) %>%
#     filter(!is.na(files)) %>%
#     mutate(
#       sheets = pmap(
#         list(data, cruise_id, files),
#         function(.x, .y, .z) {
#           tryCatch(
#           {
#               
#               update_proc_log(x = .x, cruise = .y, prev_log = .z, 
#                               type = "meta")
#               
#           },
#           error = function(e) {
#             cli_alert_danger("Cruise has issue unnesting!")
#             cli_alert_danger(sprintf(
#               "Error in %s: %s",
#               deparse(e[["call"]]),
#               e[["message"]]
#             ))
#             NULL
#           }
#           )
#         }
#       )
#     )
# # .Last.value$sheets[[1]]
# 
# openXL(.Last.value$sheets[[1]])
# 
# meta_all %>%
#     nest(.by = cruise_id) %>%
#     filter(str_detect(cruise_id, str_c(cruise_to_update, collapse = "|"))) %>%
#     mutate(files = dir_ls(here(), regexp = "WS212")) %$% 
# map2(
#   data, cruise_id,
#   (\(x, y) process_log_sheet(
#   .meta = x, cal = cal, cruise = y) %>%
#    openXL()
#   )
# ) 
```

# ---- Don't Need to Go Beyond ----

Styles
```{r old-code}
# original way to load data
# meta_all <- 
#   dir_ls(here(cloud_dir), 
#         regexp = "imars_inventory_chl_hplc_cdom") %>%
#   readxl::read_xlsx(sheet = "sample_meta_data",
#                     col_types = c(rep("guess", 14), 
#                                   "numeric", "text", "numeric")) %>%
#   
#   # remove cruises fully completed pre-COVID
#   filter(
#     year >= 2019 & !str_detect(cruise_id, c("WS19028|WS19119|WS19210"))
#     & !(cruise_id %in% crs_log_exst$cruise_id)
#     ) %>%
#       mutate(date_mm_dd_yy = as_date(date_mm_dd_yy),
#              sample_number = as.character(sample_number))


# sstyles <-
#   list(
#       styl_date     = createStyle(numFmt = "DATE"),
#       styl_time     = createStyle(numFmt = "hh:mm"),
#       styl_col1     = createStyle(fgFill = "#FDE9D9"),
#       styl_col2     = createStyle(fgFill = "#DCE6F1"),
#       styl_col3     = createStyle(fontName        = "Arial",
#                                    fontSize       = 22,
#                                    fontColour     = "white",
#                                    textDecoration = "bold",
#                                    border         = "TopBottomLeftRight",
#                                    fgFill         = "#948A54",
#                                    halign         = "center"),
#       styl_col4     = createStyle(
#                                    fgFill = "#DDD9C4",
#                                    halign = "center"),
#       styl_border   = createStyle(border       = "Right", 
#                                    borderStyle = "thin"),
#       styl_brd_bot  = createStyle(border       = "bottom", 
#                                    borderStyle = "thin"),
#       styl_brd_head = createStyle(border          = "bottom", 
#                                    borderStyle    = "thin",
#                                    textDecoration = "bold",
#                                    halign         = "center",
#                                    wrapText       = TRUE),
#       styl_align     = createStyle(halign = "center"),
#       styl_brd_dash  = createStyle(border      = "right", 
#                                    borderStyle = "dotted")
# ) 



# Metadata Sheet Formatting ---
## Read Info

# meta <-
#   meta_all %>%
#   filter(str_detect(cruise_id, cruise)) %>%
#   select(-1, -sample_number) %>%
#   mutate(sample_collection_time_gmt = hms::as_hms(sample_collection_time_gmt)) %>%
#   filter(!str_detect(sample_type, "HPLC")) %>%
#   select(
#     date_mm_dd_yy,                                
#     sample_collection_time_gmt,
#     lat,
#     lon,
#     max_depth,
#     depth_m,
#     station,
#     identifier,
#     sample_type) %>%
# 
#   pivot_wider(
#     names_from   = sample_type, 
#     values_from  = identifier,
#     values_fn    = list,
#     names_repair = janitor::make_clean_names
#     ) %>%
#   unnest(c(chl_a, cdom)) %>%
#   group_by(station, depth_m) %>%
#   mutate(cdom = replace(cdom, duplicated(cdom), NA)) %>%
#   ungroup() %>%
#   transmute(
#     "Date (UTC)"         = date_mm_dd_yy,                                
#     "Time (UTC)"         = sample_collection_time_gmt,
#     "Latitude (deg. N)"  = lat,
#     "Longitude (deg. W)" = lon,
#     "Bottom Depth (m)"   = max_depth,
#     "Sample Depth (m)"   = depth_m,
#     "Station ID"         = station,
#     "Instrument"         = "PElam850+",
#     "Sample ID"          = chl_a,
#     "CDOM Sample ID"     = cdom
#     ) %>%
#   mutate(
#     "ap,d(λ) process date"                       = NA_Date_,
#     "Instrument"                                 =  "PElam850+",
#     "ap(λ) filename (*.Sample.Raw*)"             = NA_character_,
#     "ad(λ) filename (*.Sample.Raw*)"             = NA_character_,
#     "CDOM process date"                          = NA_Date_,
#     "Instrument "                                =  "PElam850+",
#     "aCDOM(λ) filename, Rep. #1 (*.Sample.Raw*)" = NA_character_,
#     "aCDOM(λ) filename, Rep. #2 (*.Sample.Raw*)" = NA_character_,
#     "aCDOM(λ) filename, Rep. #3 (*.Sample.Raw*)" = NA_character_,
#     "Comments"                                   = NA_character_
#   ) %>%
#   relocate("ap,d(λ) process date", 
#            .before = Instrument) %>%
#   relocate("CDOM Sample ID", 
#            .after = "Instrument ")
# 
# meta


## Create Workbook

# # ---- Create Worksheet ---
# wb <- createWorkbook(
#                creator = "Sebastian Di Geronimo", 
#                title = glue("{cruise} Process Log"))
# 
# # ---- add meta data ---
# addWorksheet(wb, sheetName = sht_nm[1])
# 
# writeData(wb, sheet = sht_nm[1], startCol = 1, startRow = 1, x = meta,
#           headerStyle = sstyles$styl_brd_head)
# 
# writeData(wb, sheet = sht_nm[1], startCol = "H", startRow = 1, x = meta[,8:12],
#           headerStyle = sstyles$styl_col1)
# 
# writeData(wb, sheet = sht_nm[1], startCol = "M", startRow = 1, x = meta[,13:18],
#           headerStyle = sstyles$styl_col2)
# 
# writeData(wb, sheet = sht_nm[1], startCol = "S", startRow = 1, x = meta[, 19],
#           headerStyle = sstyles$styl_brd_head)
# 
# setRowHeights(wb, sheet = sht_nm[1], rows = 1, heights = 73.2)
# setColWidths(wb, sheet = sht_nm[1], cols = "S", widths = 39.56)
# 
# # ---- style ---
# options(openxlsx.dateFormat = "mm/dd/yy")
# addStyle(wb, sht_nm[1],
#          style      = sstyles$styl_date,
#          rows       = 1:(nrow(meta) + 1),
#          cols       = 1,
#          gridExpand = TRUE,
#          stack      = TRUE)
# 
# addStyle(wb, sht_nm[1],
#          style      = sstyles$styl_time,
#          rows       = 1:nrow(meta) + 1,
#          cols       = 2,
#          gridExpand = TRUE,
#          stack      = TRUE)
# 
# 
# addStyle(wb, 
#          sheet      = sht_nm[1], 
#          style      = sstyles$styl_col1,
#          cols       = 8:12,
#          rows       = 1:nrow(meta) + 1,
#          gridExpand = TRUE,
#          stack      = TRUE)
# 
# addStyle(wb, 
#          sheet      = sht_nm[1], 
#          style      = sstyles$styl_col2,
#          cols       = 13:18,
#          rows       = 1:nrow(meta) + 1,
#          gridExpand = TRUE,
#          stack      = TRUE)
# 
# addStyle(wb, 
#          sheet      = sht_nm[1], 
#          style      = sstyles$styl_border,
#          cols       = c("G", "L", "S", "R"),
#          rows       = 1:(nrow(meta) + 1),
#          gridExpand = TRUE, 
#          stack      = TRUE)
# 
# addStyle(wb, 
#          sheet      = sht_nm[1], 
#          style      = sstyles$styl_brd_bot,
#          cols       = 1:19,
#          rows       = nrow(meta) + 1,
#          gridExpand = TRUE, 
#          stack      = TRUE)
# 
# addStyle(wb, 
#          sheet      = sht_nm[1], 
#          style      = sstyles$styl_align,
#          cols       = 1:19,
#          rows       = 1:nrow(meta) + 1,
#          gridExpand = TRUE, 
#          stack      = TRUE)


# ---- Chl-a Process Log ---

# chl_a <-
#   meta_all %>%
#   filter(str_detect(cruise_id, cruise)) %>%
#   select(-1, -sample_number) %>%
#   mutate(sample_collection_time_gmt = hms::as_hms(sample_collection_time_gmt)) %>%
#   filter(str_detect(sample_type, "Chl-a")) %>%
#   select(
#     date_mm_dd_yy, sample_collection_time_gmt, lat, lon, max_depth, depth_m, 
#     station, identifier, vol_ml)  %>%
#   transmute(
#     "Date (UTC)"         = date_mm_dd_yy,                                
#     "Time (UTC)"         = sample_collection_time_gmt,
#     "Latitude (deg. N)"  = lat,
#     "Longitude (deg. W)" = lon,
#     "Bottom Depth (m)"   = max_depth,
#     "Sample Depth (m)"   = depth_m,
#     "Station ID"         = station,
#     "Sample ID"          = identifier,
#     "Volume seawater (ml)" = vol_ml
#     ) %>%
#   mutate(
#     "Process date"                 = NA_Date_,
#     "Volume methanol (ml)"         = NA_real_,
#     "'No Acid' Rb"                 = NA_real_,
#     "Fluorometer Calibration date" = cal$cal_date,
#     "No Acid' blank"               = NA_real_,
#     "'No Acid' Slope"              = cal$no_acid_slop,
#     "'No Acid' y-int"              = cal$no_acid_y_int,
#     "Cal. 'No Acid' Low"           = cal$cal_no_acid_low, 
#     "'No Acid' Low"                = NA_real_,               
#     "Cal. 'No Acid' High"          = cal$cal_no_acid_high,
#     "'No Acid' High"               = NA_real_,
#     "'No Acid' ratio"              = NA_real_,
#     "[Chl a] (mg m^-3)"            = NA_real_, 
#     "AVG [Chl a] (mg m^-3)"        = NA_real_, 
#     "STDEV [Chl a] (mg m^-3)"      = NA_real_,
#     "Comments"                     = NA_character_
# ) %>%
#   relocate("Process date", .after = "Sample ID")
# 


## -- Formulas for Chlorophyll-a --

# # =IF(R4>0,((Q4/R4)+(S4/T4))/2,-999)
# no_acid_ratio <- paste0(
#   "IF(", paste0("R", 1:nrow(chl_a) + 3), ">0,((", paste0("Q", 1:nrow(chl_a) + 3), 
#   "/", paste0("R", 1:nrow(chl_a) + 3), ")+(", paste0("S", 1:nrow(chl_a) + 3), 
#   "/", paste0("T", 1:nrow(chl_a) + 3), "))/2,-999)")
# 
# # V
# # =+IF(L4>0,(O4*((L4-N4)*U4)+P4)*K4/J4,-999)
# chla_conc <- 
#   paste0("+IF(", paste0("L", 1:nrow(chl_a) + 3),">0,(", 
#          paste0("O", 1:nrow(chl_a) + 3),"*((",paste0("L", 1:nrow(chl_a) + 3),
#          "-",paste0("N", 1:nrow(chl_a) + 3),")*",paste0("U", 1:nrow(chl_a) + 3),
#          ")+",paste0("P", 1:nrow(chl_a) + 3),")*",paste0("K", 1:nrow(chl_a) + 3),
#          "/",paste0("J", 1:nrow(chl_a) + 3),",-999)")
# 
# formulas <- 
# tibble(
#   # W
#   # =AVERAGE(V4:V5)
#   avg = paste0("AVERAGE(",paste0("V", seq(1, (nrow(chl_a)), 2) + 3), ":",
#                paste0("V", seq(1, (nrow(chl_a)), 2) + 4),")"),
#   # X
#   # =STDEV(V4:V5)
#   stdev = paste0("STDEV(",
#                  paste0("V", seq(1, (nrow(chl_a)), 2) + 3), ":", 
#                  paste0("V", seq(1, (nrow(chl_a)), 2) + 4),")"),
#   )  %>%
#   bind_rows(tibble(avg = rep(NA, nrow(chl_a) / 2))) %>%
#   mutate(rows = rep(1:(nrow(chl_a) / 2), 2)) %>%
#   arrange(rows, !is.na(avg), avg) %>%
#   select(-rows) %>%
#   mutate( 
#     # col U 
#     no_acid_ratio = no_acid_ratio,
#     # col V
#     chla_conc = chla_conc) 
# 
# rm(no_acid_ratio, chla_conc)


## -- Chlor-a Sheet --

# # comments
# comm <- 
#   list(
#     fluor     = createComment(
#       "*Jen*: Fluoresence Reading of the Sample (RFU)",
#       author  = "Jen", visible = FALSE),
#     blank     = createComment(
#       "*Jen*: Fluorescence reading for 100% methanol 'blank'",
#       author  = "Jen", visible = FALSE),
#     low_read  = createComment(
#       "*Jen*: Report average of all 'low' readings collected that day.",
#       author  = "Jen", visible = FALSE),
#     high_read = createComment(
#       "*Jen*: Report average of all 'high' readings collected that day.", 
#       author  = "Jen", visible = FALSE)
#     )
# # --- Create Worksheet ---
# # wb <- loadWorkbook(file_examp)
# addWorksheet(wb, sheetName = sht_nm[2])
# 
# # large header name
# mergeCells(wb, sht_nm[2], cols = 9:24, rows = 1)
# writeData(wb, sht_nm[2], "Trilogy Fluorometer", startCol = 9, startRow = 1, 
#           colNames = FALSE)
# addStyle(wb, sht_nm[2], rows = 1, cols = 9:24, stack = TRUE, style = sstyles$styl_col3)
# 
# # small header name
# mergeCells(wb, sht_nm[2], cols = 9:12,  rows = 2)
# mergeCells(wb, sht_nm[2], cols = 13:16, rows = 2)
# mergeCells(wb, sht_nm[2], cols = 17:21, rows = 2)
# mergeCells(wb, sht_nm[2], cols = 22:24, rows = 2)
# 
# 						
# 
# writeData(wb, sht_nm[2], "Secondary Solid Standard", 
#           startCol = 17, startRow = 2, colNames = FALSE)
# writeData(wb, sht_nm[2], "Welschmeyer (no acid)", 
#           startCol = 22, startRow = 2, colNames = FALSE)
# 
# 
# # data
# writeData(wb, sheet = sht_nm[2], startCol = 1, startRow = 3, x = chl_a,
#           headerStyle = sstyles$styl_brd_head)
# 
# options(openxlsx.dateFormat = "mm/dd/yy")
# addStyle(wb, sht_nm[2],
#          style      = sstyles$styl_date,
#          rows       = 3:(nrow(meta) + 3),
#          cols       = 1,
#          gridExpand = TRUE,
#          stack      = TRUE)
# 
# addStyle(wb, sht_nm[2],
#          style      = sstyles$styl_date,
#          rows       = 3:(nrow(meta) + 3),
#          cols       = "M",
#          gridExpand = TRUE,
#          stack      = TRUE)
# 
# addStyle(wb, sht_nm[2],
#          style      = sstyles$styl_time,
#          rows       = 3:(nrow(meta) + 3),
#          cols       = 2,
#          gridExpand = TRUE,
#          stack      = TRUE)
# 
# 
# 
# addStyle(wb, sht_nm[2], rows = 2:(nrow(chl_a) + 3), cols = 9:24, stack = TRUE,
#          style = sstyles$styl_col4, gridExpand = TRUE)
# 
# addStyle(wb, sht_nm[2], rows = 2:(nrow(chl_a) + 3), 
#          cols = c("H", "L", "P", "U", "X", "Y"), 
#          stack = TRUE,
#          style = sstyles$styl_border, 
#          gridExpand = TRUE)
# addStyle(wb, sht_nm[2], rows = 3:(nrow(chl_a) + 3), 
#          cols = c("T"), 
#          stack = TRUE,
#          style = sstyles$styl_brd_dash, 
#          gridExpand = TRUE)
# 
# addStyle(wb, sht_nm[2], cols = 1:ncol(chl_a),rows = nrow(chl_a) + 3, style = sstyles$styl_brd_bot,
#          stack = TRUE)
# 
# # set col/row size
# setRowHeights(wb, sheet = sht_nm[2], rows = 1, heights = 29.3)
# setRowHeights(wb, sheet = sht_nm[2], rows = 2, heights = 14.4)
# setRowHeights(wb, sheet = sht_nm[2], rows = 3, heights = 57)
# setColWidths(wb, sheet = sht_nm[2], cols = "Y", widths = 40.56)
# 
# # input formulas
# writeFormula(wb, sht_nm[2],
#              formulas$no_acid_ratio, 
#              startCol = "U",
#              startRow = 4)
# writeFormula(wb, sht_nm[2],
#              formulas$chla_conc, 
#              startCol = "V",
#              startRow = 4)
# writeFormula(wb, sht_nm[2],
#              formulas$avg, 
#              startCol = "W",
#              startRow = 4)
# writeFormula(wb, sht_nm[2],
#              formulas$stdev, 
#              startCol = "X",
#              startRow = 4)
# 
# # write comments
# writeComment(wb, sht_nm[2], col = "L", row = 4, comment = comm$fluor)
# writeComment(wb, sht_nm[2], col = "N", row = 4, comment = comm$blank)
# writeComment(wb, sht_nm[2], col = "R", row = 4, comment = comm$low_read)
# writeComment(wb, sht_nm[2], col = "T", row = 4, comment = comm$high_read)


```




